```{r global_options, echo=FALSE, results='hide'}
library(knitr)
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.align='center', results='hide', echo=FALSE, warning=FALSE, message=FALSE, dev=c('pdf', 'png'), cache=FALSE)
```

```{r}
knitr::opts_knit$set(base.dir='/Users/Moana/Dropbox/Moana_Time_Series/Manuscript #1_Surface Time series/Ecotones paper/R_analysis/', fig.path='/Users/Moana/Dropbox/Moana_Time_Series/Manuscript #1_Surface Time series/Ecotones paper/R_analysis/Figures_10000/')
```


```{r Load libraries and data}
library(phyloseq)  
library(ggplot2)
library(plyr)
library(dplyr)
library(scales)
library(reshape)
library(reshape2)
library(RColorBrewer)
library(grid)

uzdir <- "/Users/Moana/Dropbox/Moana_Time_Series/Manuscript #1_Surface Time series/Ecotones paper/R_analysis/"
otutable_biom_file <- paste("/Users/Moana/Dropbox/Moana_Time_Series/Manuscript #1_Surface Time series/Ecotones paper/R_analysis/", "10000_merged_otu_table_json.biom", sep = "")
map_file <- paste("/Users/Moana/Dropbox/Moana_Time_Series/Manuscript #1_Surface Time series/Ecotones paper/R_analysis/", "Ecotone_map.txt", sep = "")

```

```{r Create phyloseq object}

# Now import the .biom-formatted otu_table-tax_table file.

biom_otu_tax <- import_biom(otutable_biom_file)

# Add sample data to the dataset using merge
bmsd <- import_qiime_sample_data(map_file)
class(bmsd)
dim(bmsd)
sample_data(bmsd)
biom_otu_tax


#Merge into phyloseq format
MM_phyloseq <- merge_phyloseq(biom_otu_tax, bmsd)
MM_phyloseq
sample_sums(MM_phyloseq)

#Save original phyloseq file
MM_phyloseq0 = MM_phyloseq
```

#####Create average result for multiple rarefaction by transforming data using (divide by 10) and check counts per sample
```{r Create average result for multiple rarefaction by transforming data using (divide by 10), results='markup'}
MM_phyloseq = transform_sample_counts(MM_phyloseq, function(x) x/10)
sample_sums(MM_phyloseq)
```

##### Round and confirm count number
```{r Round and confirm count number, results='markup'}
MM_phyloseq = transform_sample_counts(MM_phyloseq, round)
sample_sums(MM_phyloseq)
MM_phyloseq = prune_samples(sample_sums(MM_phyloseq)>=1, MM_phyloseq)
sample_sums(MM_phyloseq)


#check that all numbers are integers
otu_table_df<- as.data.frame(otu_table(MM_phyloseq))
```

Check that all OTUs have representative counts  
For here taxa = OTU  
__Commands interpretation:__  
_Total number of taxa in dataset:_ sum(taxa_sums(MM_phyloseq) > 0)   

_Any taxa with no hits:_ any(taxa_sums(MM_phyloseq)== 0)

```{r identify taxa with only zeros, results='markup', echo=TRUE}
sum(taxa_sums(MM_phyloseq) > 0)
any(taxa_sums(MM_phyloseq)== 0)
sum(taxa_sums(MM_phyloseq) == 0)
any(taxa_sums(MM_phyloseq) > 1)
sum(taxa_sums(MM_phyloseq) > 1)
any(taxa_sums(MM_phyloseq) < 1)
sum(taxa_sums(MM_phyloseq) < 1)
```

#####Prune taxa with less than 1 count and check taxa numbers again
```{r  Save original file and create new file with only present (no zeroes) taxa, results='markup', echo=TRUE}

#Create new file with only present (no zeroes) taxa

MM_phyloseq = prune_taxa(taxa_sums(MM_phyloseq) > 1, MM_phyloseq)
any(sample_sums(MM_phyloseq) == 0)
any(sample_sums(MM_phyloseq) > 0)
sum(taxa_sums(MM_phyloseq) > 0)
any(sample_sums(MM_phyloseq) < 1)
sum(taxa_sums(MM_phyloseq) < 1)
```

###Compare sequences per sample or OTU
```{r Compare sequences per sample or OTU}
readsumsdf = data.frame(nreads = sort(taxa_sums(MM_phyloseq), TRUE), sorted = 1:ntaxa(MM_phyloseq), 
                        type = "OTU")
readsumsdf = rbind(readsumsdf, data.frame(nreads = sort(sample_sums(MM_phyloseq), 
                                                        TRUE), sorted = 1:nsamples(MM_phyloseq), type = "Samples"))
title = "Total number of reads"
p = ggplot(readsumsdf, aes(x = sorted, y = nreads)) + geom_bar(stat = "identity")
p + ggtitle(title) + scale_y_log10() + facet_wrap(~type, 1, scales = "free")
```

```{r}
sample_variables(MM_phyloseq)
```

```{r Attached OTU ID}
tax_table(MM_phyloseq) <- cbind(tax_table(MM_phyloseq), OTU=taxa_names(MM_phyloseq))
```

```{r Rename Ranks}
colnames(tax_table(MM_phyloseq)) = c("Domain", "Phylum", "Class", "Order", "Family", "Genus", "Species", "OTU")
tax_table(MM_phyloseq) =gsub("D_1__", "", tax_table(MM_phyloseq))
```

# Subset samples and only keep surface samples
```{r Subset for surface}
MM_surf_10000_phyloseq = subset_samples(MM_phyloseq, Depth == "0")
sample_sums(MM_surf_10000_phyloseq)
MM_surf_10000_phyloseq
```

```{r Calculate Alpha diversity}
alpha_summary <- estimate_richness(MM_surf_10000_phyloseq, measures = c("Observed", "Shannon"))
#combine with metadata
alpha_meta <- data.frame(alpha_summary, sample_data(MM_surf_10000_phyloseq))
```

```{r Summarize richness data and calcuate SE SD CI mean}
library(Rmisc)
# mean will be calculated from the Observed richness column by all the variables listed in groupvars.
tgc_Rich <- summarySE(alpha_meta, measurevar="Observed", groupvars=c("Month","Year", "Time", "Water_mass_PCA"))
tgc_Rich

# repeat for Shannon
tgc_S <- summarySE(alpha_meta, measurevar="Shannon", groupvars=c("Month","Year", "Time", "Water_mass_PCA"))
tgc_S


#rename headers to differentiate results for Shannon and Observed
colnames(tgc_Rich) = c("Month", "Year", "Time", "Water_mass_PCA", "N_Rich", "Observed", "sd_observed", "se_observed", "ci_observed")

colnames(tgc_S) = c("Month", "Year", "Time", "Water_mass_PCA", "N_Shannon", "Shannon", "sd_shannon", "se_shannon", "ci_shannon")

#combine results
tgc <- merge(tgc_S,tgc_Rich,by=c("Month", "Year", "Time", "Water_mass_PCA"))

tgc$Water_mass_PCA <- factor(tgc$Water_mass_PCA,
                             levels = c("NW", "STW", "FRONT", "SAW"))

tgc$Month <- factor(tgc$Month,
                    levels = c("January", "June", "July", "December", "March", "April"))

```

##########################################################################################################################

#Make OTU data frame
```{r}
library(ggplot2)
library(plyr)
library(scales)
library(reshape)
library(RColorBrewer)
library(grid)
library(phyloseq)

speciesList <- tapply(sample_names(MM_surf_10000_phyloseq), get_variable(MM_surf_10000_phyloseq, "SampleID"), c) #Makes columns of water mass types and says which samles are found there
speciesPhyseq <- lapply(speciesList, prune_samples, MM_surf_10000_phyloseq) #only samples that have these species will be kept
speciesOTUtable <- lapply(speciesPhyseq,otu_table) #go through data and extract the information that matches that data
speciesAvg <- lapply(speciesOTUtable,rowMeans) #what does each species contribute to the 100%? percent transformation
pooledOTUtable = t(do.call(rbind,speciesAvg)) #Take the new average and bind it to original document
#This is as far back as I could trace the x
pooledOTUtable = data.frame(OTU=row.names(pooledOTUtable),pooledOTUtable) #converts data to a dataframe (note that this is subsetted!!)
TT = tax_table(MM_surf_10000_phyloseq) #attaches taxonomy to those species
TT = TT[, which(apply(!apply(TT, 2, is.na), 2, any))]
tdf = data.frame(TT, OTU = taxa_names(MM_surf_10000_phyloseq)) #attaches OTUs to data frame
pOTUtax = merge(pooledOTUtable, tdf, by.x = "OTU")
pOTU = data.frame(pOTUtax,SeqTotal = rowSums(pOTUtax[,2:85])) #numbers show in which columns treatment data is located
pOTU.phylum =pOTU[,c(2:85,1)] #need to change this and the above numbers manually. The 1 refers to the column of taxonomy data
melt.phylum = melt(pOTU.phylum,id.vars="OTU") #melt all the OTUs that are one particular thing into a rank. Still has many observations for each OTU in each water mass
colnames(melt.phylum)[2]="SampleID"
agg.phylum=aggregate(.~OTU+SampleID,melt.phylum,sum) #Turns it into percentages
agg.phylum$SampleID = gsub( "X", "", paste(agg.phylum$SampleID)) ##########THIS IS MY FIX!!!!!!#######################

library(reshape2)
data = dcast(agg.phylum, SampleID ~ OTU) #Rearrages so that each species has a column


PhylumCount = data
# Metadata = bmsd

# write.csv(bmsd, file = "C://Users//Moana//Documents//Uni//2015//ECOL491//All data//data//bmsdMetadata.csv", row.names=FALSE)
# write.csv(data, file = "C://Users//Moana//Documents//Uni//2015//ECOL491//All data//data//dataPhyla.csv", row.names=FALSE)

Metadata = read.table("C://Users/Moana/Documents/Uni/2016/Publication/Data from Rachel/Ecotone_map.txt", header = TRUE)

# PhylumCount = read.csv("C://Users//Moana//Documents//Uni//2015//ECOL491//All data//data//dataPhyla.csv", header = TRUE)

NewDF <- merge(PhylumCount, Metadata, by="SampleID")#, all.x=TRUE, all.y=TRUE)
# NewDF$Chlorophylla <- NULL
# NewDF$BA <- NULL
# NewDF$BP <- NULL
# NewDF <- na.omit(NewDF)

#write.csv(NewDF, file = "C://Users//Moana//Documents//Uni//2015//ECOL491//All data//data//NewOTUMetadatainclNA.csv", row.names=FALSE)


PhylumCount2 = agg.phylum
NewDF2 <- merge(PhylumCount2, Metadata, by="SampleID")#, all.x=TRUE, all.y=TRUE)
# NewDF2$Chlorophylla <- NULL
# NewDF2$BA <- NULL
# NewDF2$BP <- NULL
# NewDF2 <- na.omit(NewDF2)

#write.csv(NewDF2, file = "C://Users//Moana//Documents//Uni//2015//ECOL491//All data//data//NewOTUMetadataLayout2inclNA.csv", row.names=FALSE)

NewDF3 <- merge(NewDF2, tdf, by="OTU", all.x=TRUE, all.y=TRUE) #tdf adds all the other taxonomy
#write.csv(NewDF3, file = "C:/Users/Moana/Documents/Uni/2016/Publication/paper/NewFulltaxonomyinclNA.csv", row.names=FALSE)


```